---
title: "DataHack- Upworthy"
author: "Megumi Low, Daniel Miao, Mohammed Mehboob, Jas Dinh"
date: "April 9, 2022"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

1. Did Upworthy really change their clickbait publishing practices?
2. What was the impact of that change?
3. Do you recommend an investment in Upworthy?


Read CSV files
```{r readCSV}
country_df = read.csv("country-data.csv") # aggregates user analytics by country
daily_user_df = read.csv("daily-user-info.csv") # user analytics from 2013 to 2015
archive_df = read.csv("upworthy-archive.csv") # A/B testing results
```

Import packages
```{r importPackages, results = "hide"}
library(dplyr)
library(ggplot2)
library(stringr)
```

## Initial Analysis with Generalized Linear Models

We first defined clickbait as something that is *attention-grabbing*, *emotionally arousing*, and *misleading*. Since, there are three different datasets with a total of 31 variables, we were able to consider different metrics for clickbait such as the clicks per impressions, the language of the headline, and the bounce rate of the article.

```{r newVars}
# Data Wrangling

# create numeric variable of winner for logistic regression
archive_df$bWinner <- as.numeric(as.logical(archive_df$winner))
# create clicks per impression variable
archive_df$click_imp <- archive_df$clicks / archive_df$impressions
```

### Data Analysis on Clicks per Impression
```{r glms}
click_glm = glm(bWinner ~ click_imp, data = archive_df)
summary(click_glm)

score_glm = glm(bWinner ~ score, data = archive_df)
summary(score_glm)

cor.test(archive_df$bWinner, archive_df$click_imp)
cor.test(archive_df$bWinner, archive_df$score)

ggplot(archive_df, aes(x = created_at, y = click_imp)) + geom_point(aes(color = winner)) + geom_smooth() + labs(title = "Clicks per Impressions over Time", x= "Date of A/B Test", y = "Clicks per Impressions")
```


After running a logistic regression model of winner and clicks per impression as well as score of the A/B test, the results show that there is a significant relationship between the two variables. However, the correlation is rather low and brings up the question why Upworthy did not select the article with the highest A/B testing score or clicks per impression.

## Text Mining Headlines

```{r textMining}
# using punctuation to evaluate clickbait
punctuations.dict <- c(question = "?",
                       quote = "'",
                       dollar = "$",
                       exclaim = "!")
archive_df$created_at <- as.Date(archive_df$created_at)
archive_df$headline_nchar <- archive_df$package_headline %>% nchar()
archive_df$headline_nwords <- archive_df$package_headline %>% str_count(" ") + 1

for(i in punctuations.dict %>% names()){archive_df[[i]] <- ifelse(grepl(paste0("[",punctuations.dict[i],"]"), archive_df$package_headline), T, F) }
```


### Data Visualization on Punctuation
```{r punctuationGraphs}
archive_df %>% ggplot(aes(created_at, clicks/impressions, color = dollar)) + geom_smooth() + labs(title = "Clicks with Dollar Marks in Headline", x= "Date", y= "Total Clicks Per Impressions")
archive_df %>% ggplot(aes(created_at, clicks/impressions, color = question)) + geom_smooth()+ labs(title = "Clicks with Question Marks in Headline", x= "Date", y= "Total Clicks Per Impressions")
archive_df %>% ggplot(aes(created_at, clicks/impressions, color = quote)) + geom_smooth() + labs(title = "Clicks with Quotes in Headline", x= "Date", y= "Total Clicks Per Impressions")
archive_df %>% ggplot(aes(created_at, clicks/impressions, color = exclaim)) + geom_smooth() + labs(title = "Clicks with Exclaimation Marks in Headline", x= "Date", y= "Total Clicks Per Impressions")
```


We wanted to analyze the headlines of each article in the A/B testing packages and used punctuation as a metric for clickbait. Contrary to our original belief, the punctuation marks actually garnered less clicks during A/B testing.

```{r lmClicks}
model.1 <- lm(click_imp ~ question+exclaim+dollar+quote+headline_nchar+headline_nwords, data = archive_df)
summary(model.1)
```


After running a linear model, we confirmed that the punctuation marks led to significantly less clicks per impression with the exception of number of characters in the headline.

## Natural Language Processing Sentiment Analysis

To view the NLP sentiment analysis model, view the Colab Notebook here: [https://colab.research.google.com/drive/11nhDAOWKevdZQ1J4-eo2g0YU_heuWFjF?usp=sharing](https://colab.research.google.com/drive/11nhDAOWKevdZQ1J4-eo2g0YU_heuWFjF?usp=sharing)

In summary, using pre-trained neural networks of HuggingFace, we identified clickbait using sentiment analysis models to see if a collection of text is positive or negative. 

We evaluated that the overall strength of the sentiment does not decrease, but the headlines may be getting more positive near the date of the clickbait pledge. 

In a breakdown of each emotion, we found that most articles either conveyed joy or anger. Additionally, there was an overall increase in sentiments for winner articles overtime.

## Time Series Analysis

### Loess Regression on Users
```{r loessUsers}
# Wrangle date variable and prepare loess
daily_user_df$dayX <- as.Date(daily_user_df$day,format="%m/%d/%y")
daily_user_df$index <- 1:nrow(daily_user_df)

loess_users <- loess(users ~ index, data = daily_user_df, span=0.25)
smoothed_users <- predict(loess_users)

#create Loess regression
plot(y = daily_user_df$users, x = daily_user_df$dayX, type="l", main="Loess Smoothing and Prediction on Users", xlab="Date", ylab="Users")
lines(smoothed_users, x=daily_user_df$dayX, col="blue")

```

### Loess Regression on Bounce Rate
```{r loessBounce}
loess_bounce <- loess(bounce_rate ~ index, data = daily_user_df, span=0.25)
smoothed_bounce <- predict(loess_bounce)

#create Loess regression
plot(y = daily_user_df$bounce_rate, x = daily_user_df$dayX, type="l", main="Loess Smoothing and Prediction on Bounce Rate", xlab="Date", ylab="Bounce Rate")
lines(smoothed_bounce, x=daily_user_df$dayX, col="blue")
```


After evaluating the clickbait in articles, we then investigated its impact on the company. Using a Loess Regression model, it is predicted that the bounce rate of the articles will stay consistent while the amount of users for the website will decrease.

## Conclusion

After analyzing different clickbait metrics such as clicks per impression, sentiment analysis, and bounce rate, we determined that Upworthy has not changed their clickbait publishing practices. However, the impact of the clickbait free pledge on the company has led to declining user engagement, retention, and new users. Due to the negative forecast on the website's user analytics, we do not recomment an investment in Upworthy.
